---
title: "CI Integration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CI/CD Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(cohortTdd)
```

# Automated Testing with CI

The `cohortTdd` package supports automated testing through continuous integration pipelines. This enables systematic validation of cohort definitions as they evolve.

## GitHub Actions Integration

### Automatic Setup

When creating a new project, a GitHub Actions workflow is included by default:

```{r github-actions-setup}
setup_cohort_testing_project(
  project_name = "my-cohort-test",
  cohort_name = "diabetes_cohort", 
  author_name = "Your Name",
  github_actions = TRUE  # Default
)
```

This creates `.github/workflows/test-cohorts.yml` with:

- **R environment setup** (latest R version)
- **Dependency installation** (cohortTdd and requirements)
- **Test execution** on push/pull requests
- **Results reporting** with test summaries

### Manual Workflow Configuration

For existing projects, create `.github/workflows/test-cohorts.yml`:

```yaml
name: Test Cohort Definitions

on:
  push:
    branches: [ main, develop ]
  pull_request:
    branches: [ main ]

jobs:
  test:
    runs-on: ubuntu-latest
    
    steps:
    - uses: actions/checkout@v3
    
    - uses: r-lib/actions/setup-r@v2
      with:
        r-version: 'release'
        
    - name: Install system dependencies
      run: |
        sudo apt-get update
        sudo apt-get install -y libcurl4-openssl-dev libxml2-dev
        
    - name: Install R dependencies
      run: |
        install.packages(c("remotes", "testthat"))
        remotes::install_deps(dependencies = TRUE)
        remotes::install_github("bill-baumgartner/cohortTdd")
      shell: Rscript {0}
      
    - name: Run cohort validation tests
      run: |
        library(testthat)
        library(cohortTdd)
        test_check("your-project-name")
      shell: Rscript {0}
```

## Test Structure for CI

### Automated Test Files

Structure your tests for CI compatibility:

```{r ci-test-structure}
# tests/testthat/test-cohort-validation.R
library(testthat)
library(cohortTdd)

test_that("Primary cohort validation passes", {
  # Set shorter timeout for CI
  options(timeout = 600)  # 10 minutes max
  
  results <- validate_my_cohort(
    patients_file = system.file("testdata", "patients.xlsx", package = "your.package"),
    cohort_file = system.file("testdata", "cohort.json", package = "your.package"),
    print_summary = FALSE,
    verbose = FALSE
  )
  
  expect_equal(results$failed_tests, 0)
  expect_gt(results$total_tests, 0)
  expect_equal(results$passed_tests, results$total_tests)
})
```

### Environment Variables

Configure CI environment for reproducible tests:

```{r ci-environment}
# In .github/workflows/test-cohorts.yml
env:
  EUNOMIA_DATA_FOLDER: ${{ runner.temp }}/eunomia_data
  R_REMOTES_NO_ERRORS_FROM_WARNINGS: true
  _R_CHECK_FORCE_SUGGESTS_: false
```


This CI integration ensures your cohort definitions are automatically validated with every code change, maintaining quality and catching regressions early.