---
title: "Getting Started with cohortTdd"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with cohortTdd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE  # Don't run examples during package build
)
```

```{r setup}
library(cohortTdd)
```

# Introduction

The `cohortTdd` package provides a framework for test-driven development of OHDSI cohort definitions. It allows you to easily validate cohort logic against synthetic patient data with known inclusion/exclusion patterns.

## Core Concept

Test-driven development for cohorts follows this workflow:

1. **Create test patients** with known characteristics
2. **Define expected outcomes** (included/excluded with specific reasons)
3. **Develop your cohort definition** (as ATLAS JSON)
4. **Test your cohort definition** against the synthetic data
5. **Iterate** until all tests pass

## Quick Start

### Option 1: Direct Validation

The simplest way to test a cohort:

```{r direct-validation}
# Test your cohort directly
results <- validate_my_cohort(
  patients_file = "my_patients.xlsx",
  cohort_file = "my_cohort.json"
)

# View results
print(results)
```

### Option 2: Create a Testing Project

For more complex testing with CI/CD integration:

```{r create-project}
# Create a complete testing project
setup_cohort_testing_project(
  project_name = "my-diabetes-cohort-test",
  cohort_name = "diabetes_cohort",
  author_name = "Your Name"
)
```

This creates:
- Template test files using R package conventions
- Example patient data
- Proper directory structure (`inst/testdata/`)
- GitHub Actions workflow for CI/CD
- README with instructions

## Patient Data Format

Your Excel file should contain:

### Required Tables
Standard OMOP CDM tables like:
- `person` - Patient demographics
- `observation_period` - Follow-up periods
- `drug_exposure` - Medication data
- `condition_occurrence` - Diagnosis data
- `measurement` - Lab values, vitals
- `procedure_occurrence` - Procedures

### Required Explanation Tab
A special `explanation` sheet with:

| Column | Description |
|--------|-------------|
| `patient_id` | Patient identifier |
| `reason_id` | Expected exclusion reason ID (NA for included patients) |
| `explanation` | Free text description |

The reason ID is an integer linked to the order of constraints in your cohort definition
1 = excluded based on the cohort entry events
2 = excluded based on the first inclusion criteria
3 = excluded based on the second inclusion criteria
n = excluded based on the (n-1)th inclusion criteria

### Example Structure

```{r patient-data-example}
# Example explanation tab:
explanation <- data.frame(
  patient_id = 1:4,
  reason_id = c(NA, 1, 2, NA),
  explanation = c(
    "INCLUDE - Meets all criteria",
    "EXCLUDE - Failed cohort entry criteria",
    "EXCLUDE - Failed first inclusion criteria", 
    "INCLUDE - Meets all criteria"
  )
)
```

## Performance Notes

- **Initial setup:** ~90 seconds (downloads empty CDM data)
- **Per patient:** ~5-10 seconds (cohort generation overhead)
- **Optimization:** Uses shared CDM to minimize setup time

This is a proof-of-concept; further optimization would be needed for production use with large numbers of test patients.

## Next Steps

- See `vignette("ci-integration")` for automated testing
- Check out the `cohortTdd.demo` project for a complete example